---
title: "Final Project"
author: "Ettore Falde, Samoussa Fofana, Federico Basaglia"
date: "11/29/2021"
output: pdf_document
---


# Introduction
In this report we consider that the new CEO of a specific IT company has contacted us because she wants us to **analyze the current Human Resources status** of the company. 
\ 
She has just sent a data set with all available employee information. As we can see, the **company has two locations**: the first one in **London**, and the second one in **Barcelona**.

The new CEO is concerned about several issues. She truly believes in gender equality in organizations as it implies a signal to society. On the other hand, she is concerned that the offices in Barcelona do not follow a similar structure to the one in London. **In her opinion, the structure of the Barcelona offices should tend towards the London structure.** In her meeting with us, she also told us that she would like to know the attitudes (e.g., satisfaction) of the employees across the different departments and if anything could be done to improve them. Finally, she commented that she is very concerned about the company's succession strategy and in particular some positions in certain departments.

Based on this information, we will to carry out an exploratory data analysis and prepare a technical report (with Rmarkdown) and a technical presentation (5-10 minutes).

# Setup the software

The software used for the development of the study and the writing of the report is R[1]. The first step is to define the work directory and to load the libraries needed:  

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Directory
setwd("/Users/ettorefalde/Documents/Ettore\ Falde/UNI/UPC/Tools\ for\ Decision\ Making\ /Final\ Project")
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(janitor)
library(caTools)
library(ROCR)
library(corrplot)
library(readr)
```

## Importing Data

The first step is to load the dataset in the system, and check the names of the variables. 

```{r}
mydb <- read.csv2("dataset.csv")
names (mydb)
```

We got the dataset from the website of Atenea, it is composed by 1506 observations of 36 variables.
The variables selected for this dataset  are:

1. **Age**: Variable that represent the age of the employee
2. **Attrition**: variable that represent the departure of employees from the organization for any reason
3. **BusinessTravel**: Represent how often an employee travel for work purpose 
4. **DailyRate**: The amount of money the employees are paid per day
5. **Department**: Department of the company at which the employee belong
6. **DistanceFromHome**: Employee home distance from the workplace 
7. **Education**: Educational level of the employee (1=Below College, 2=College,
3=Bachelor, 4=Master,5 = Doctor)
8. **EducationField**: Education field of employee (Human Resources, Life Sciencies, Marketing, Medical, Technical Degree, Other)
9. **EmployeeCount**: Coolumn all equal to 1 to count the total number of employee in the data set
10. **EmployeeNumber**: unique number to identify the employee
11. **EnvironmentSatisfaction**: level of environment satisfaction (1=Low, 2=Medium, 3=High, 4=Very High)
12. **Gender**: Gender of the employee (Male, Female) 
13. **HourlyRate**: The amount of money the employees are paid per hour
14. **JobInvolvement**: Level of involvement of the employee (1=Low, 2=Medium, 3=High, 4=Very High)
15. **JobLevel**: Is a category of authority in the company (1=low, 5=High)
16. **JobRole**: Represent the role cover by the employee (Sales Executive, Research Scientist, Laboratory Technician, Manufacturing Director, Healthcare Representative, Manager, Sales Representative, Research Director, Human Resources)
17. **JobSatisfaction**: Level of satisfaction of the employee (1=Low, 2=Medium, 3=High, 4=Very High)
18. **MaritalStatus**: Marital status of the employee ( Divorced, Married, Single)
19. **MonthlyIncome**: Monthly income of the employee 
20. **MonthlyRate**: Monthly rate of employee
21. **NumCompaniesWorked**: Number of companies for ehich the employee worked
22. **Over18**: If the age of the employee is higher than 18 (Y = yes, N = no)
23. **OverTime**: If the employee perform over time (Yes, No) 
24. **PercentSalaryHike**: Represent the percentage inrease of a salary
25. **PerformanceRating**: Performance rating of the employee (1=Low, 2=Good,3=Excellent,4=Outstanding)

26. **RelationshipSatisfaction**: whether or not the patient developed coronary heart disease in the next 10 years
27. **StandardHours**: whether or not the patient developed coronary heart disease in the next 10 years
28. **StockOptionLevel**: whether or not the patient developed coronary heart disease in the next 10 years
29. **TotalWorkingYears**: whether or not the patient developed coronary heart disease in the next 10 years
30. **TrainingTimesLastYear**: whether or not the patient developed coronary heart disease in the next 10 years
31. **WorkLifeBalance**: whether or not the patient developed coronary heart disease in the next 10 years
32. **YearsAtCompany**: whether or not the patient developed coronary heart disease in the next 10 years
33. **YearsInCurrentRole**: whether or not the patient developed coronary heart disease in the next 10 years
34. **YearsSinceLastPromotion**: whether or not the patient developed coronary heart disease in the next 10 years
35. **YearsWithCurrManager**: whether or not the patient developed coronary heart disease in the next 10 years
36. **City**: whether or not the patient developed coronary heart disease in the next 10 years


# Cleaning Data
...


# Analysis
...

# Conclusions
...

# Referemces
1. _M., L. (2004). Moneyball: The art of winning an unfair game. New york: John Wiley and sons._
2. Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
3. H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
4. David Robinson, Alex Hayes and Simon Couch (2021). broom: Convert Statistical Objects into Tidy Tibbles. R package version 0.7.9. https://CRAN.R-project.org/package=broom



