---
title: "Final Project"
author: "Ettore Falde, Samoussa Fofana, Federico Basaglia"
date: "11/29/2021"
output: pdf_document
---


# Introduction
...


# Setup the software

The software used for the development of the study and the writing of the report is R[1]. The first step is to define the work directory and to load the libraries needed:  

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Directory
setwd("/Users/ettorefalde/Documents/Ettore\ Falde/UNI/UPC/Tools\ for\ Decision\ Making\ /Final\ Project")
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(caTools)
library(ROCR)
library(corrplot)
```

## Importing Data

The first step is to load the dataset in the system, and check the names of the variables. 

```{r}
mydb <- read.csv2("dataset.csv")
names (mydb)
```

We got the dataset from the website of Atenea.
The variables selected for this dataset  are:

1. **Age**:  boolean variable, male = 1, female = 0
2. **Attrition**: age of the person
3. **BusinessTravel**: level of education 
4. **DailyRate**: boolean variable, smoker = 1, non-smoker = 0
5. **Department**: number of sigarettes smoked in one day
6. **DistanceFromHome**: medical history risk factors
7. **Education**: boolean var, predisposition to get stroke
8. **EducationField**: boolean var, predisposition to get hyp
9. **EmployeeCount**: boolean var, predisposition to get diabetes
10. **EmployeeNumber**:  total cholesterol level
11. **EnvironmentSatisfaction**: systolic blood pressure
12. **Gender**: diastolic blood pressure 
13. **HourlyRate**: Body Mass Index
14. **JobInvolvement**:  heartRate
15. **JobLevel**: glucose level
16. **JobRole**: whether or not the patient developed coronary heart disease in the next 10 years
17. **JobSatisfaction**: whether or not the patient developed coronary heart disease in the next 10 years
18. **MaritalStatus**: whether or not the patient developed coronary heart disease in the next 10 years
19. **MonthlyIncome**: whether or not the patient developed coronary heart disease in the next 10 years
20. **MonthlyRate**: whether or not the patient developed coronary heart disease in the next 10 years
21. **NumCompaniesWorked**: whether or not the patient developed coronary heart disease in the next 10 years
22. **Over18**: whether or not the patient developed coronary heart disease in the next 10 years
23. **OverTime**: whether or not the patient developed coronary heart disease in the next 10 years
24. **PercentSalaryHike**: whether or not the patient developed coronary heart disease in the next 10 years
25. **PerformanceRating**: whether or not the patient developed coronary heart disease in the next 10 years
26. **RelationshipSatisfaction**: whether or not the patient developed coronary heart disease in the next 10 years
27. **StandardHours**: whether or not the patient developed coronary heart disease in the next 10 years
28. **StockOptionLevel**: whether or not the patient developed coronary heart disease in the next 10 years
29. **TotalWorkingYears**: whether or not the patient developed coronary heart disease in the next 10 years
30. **TrainingTimesLastYear**: whether or not the patient developed coronary heart disease in the next 10 years
31. **WorkLifeBalance**: whether or not the patient developed coronary heart disease in the next 10 years
32. **YearsAtCompany**: whether or not the patient developed coronary heart disease in the next 10 years
33. **YearsInCurrentRole**: whether or not the patient developed coronary heart disease in the next 10 years
34. **YearsSinceLastPromotion**: whether or not the patient developed coronary heart disease in the next 10 years
35. **YearsWithCurrManager**: whether or not the patient developed coronary heart disease in the next 10 years
36. **City**: whether or not the patient developed coronary heart disease in the next 10 years


# Cleaning Data
...


# Analysis
...

# Conclusions
...

# Referemces
1. _M., L. (2004). Moneyball: The art of winning an unfair game. New york: John Wiley and sons._
2. Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
3. H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
4. David Robinson, Alex Hayes and Simon Couch (2021). broom: Convert Statistical Objects into Tidy Tibbles. R package version 0.7.9. https://CRAN.R-project.org/package=broom



